---
title: "Vaccine repertoire data information"
output:
  html_notebook:
    toc: yes
    toc_depth: 3
    toc_float: yes

---

```{r, message=F}
library(tidyverse)
```

```{r}
parse_catalog_tags <- function(tags){
  tag_list <- str_split(unlist(str_split(tags,",")), ":")
  tags <- sapply(tag_list, function(x) x[1])
  values <- sapply(tag_list, function(x) x[2])
  data.frame(tag = tags, value = values) %>% 
    group_by(tag) %>% 
    summarise(value = paste(value, collapse = "_"))
}
```


```{r, message = F}
df <- read_csv("/labs/khatrilab/solomonb/immune_access/compiled_sample_metadata.csv") 
```

```{r}
df_meta <- df %>% 
  select(project, sample_name, sample_catalog_tags) %>% 
  drop_na() %>% 
  mutate(parse_tags = map(sample_catalog_tags, parse_catalog_tags)) %>% 
  select(-sample_catalog_tags) %>% 
  unnest(parse_tags) %>% 
  pivot_wider(names_from = "tag", values_from = "value")
df_meta
```


# Background

# Samples

### SARS-CoV-2

- SARS-CoV-2 vaccine - 122 subjects pre- and post-vaccination - 51 - Swanson
- SARS-CoV-2 vaccine 
  - Before and after vaccination in 10 COVID-19 convalescent patients and 4 SARS-CoV-2 naïve healthy donor vaccine recipients
  - 107 total samples
  - https://www.sciencedirect.com/science/article/pii/S2352396422002298?via%3Dihub
    - 4 Healthy vaccine recipients (VRs)
      - VR8 had two Pfizer-BioNTech (BNT162b2) vaccinations prior to boost and VR10 had two Moderna (mRNA-1273) vaccinations prior to boost
    - 10 COVID-19 convalescent patients (CCPs)
      -  Pfizer-BioNTech (BNT162b2; n = 8) or Moderna (mRNA-1273; n = 2)
      - CCP2 has well controlled HIV on antiretroviral therapy
      - Post-COVID samples were obtained from CCPs a median of 101 days after COVID-19 diagnosis 
      - Post vaccine samples were taken from CCPs and VRs a median of 72 days after the second dose of the vaccine
      - CCP6 and CCP11 were sampled immediately prior to vaccination
      - CCP6, CCP11, and CCP16 were sampled both early post vaccination (day +14) and late post vaccination (day +241, +227, and +161, respectively)
    
  - Project ID: `dykema-2022-ebiomed`
  - Data ID: `4ae9ad1c-de36-4acf-80ce-9e799d79abe0`
  
```{r, message=F}
read_tsv("/labs/khatrilab/solomonb/immune_access/data/4ae9ad1c-de36-4acf-80ce-9e799d79abe0/samples.tsv") %>% 
  select(sample_name, sample_tags) %>% 
  separate(sample_name, into = c("name1", "name3", "name4"), sep = "_", remove = F) %>% 
  mutate(name2 = str_extract(name1, "Pre|Post"), name1 = gsub("Pre|Post","",name1)) %>% 
  select(sample_name, contains("name"), sample_tags)
```
  - 
    - `name1` only contains: `CCP6, VR10, VR14, VR8`, seems to be missing samples?
    - Not sure what `name3` represents: `117, 146, 229ES, COV2S, HIV, NL63S, NP, OC43S`



- SARS-CoV-2 vaccine - 
  - 20 Ad26.COV2.S vaccinated individuals, 5 placebo, 8 convalescent
  - Vx and placebo collected at d63
  - TCRb
  - https://www.nature.com/articles/s41586-021-03681-2
  - Project ID: `alter-2021-n`
  - Data ID: `322a46d5-7ed3-4749-9799-f0df1fe6d0a`
  
```{r, message=F}
read_tsv("/labs/khatrilab/solomonb/immune_access/data/322a46d5-7ed3-4749-9799-f0df1fe6d0a/samples.tsv")
```

### Influenza  
  
- Influenza vaccine - 40 individuals pre- and post-vaccination
  - cTfh - circulating CD4 cells similar to Tfhs. Identified population of ICOS+CD38+ cTfh, strongly induced by inflz, included hemagglutinin-tetramer+. Found vaccine response persists in the ICOS-CD38- cTfh population 
  - 102 samples
  - https://doi.org/10.1126/sciimmunol.aag2152
  - Project ID: `herati-2017-sciimmunol`
  - Data ID: `3a2ffef9-9bf6-486f-aa68-fa8c92893c5e`
```{r, message=F}
df <- read_tsv("/labs/khatrilab/solomonb/immune_access/data/3a2ffef9-9bf6-486f-aa68-fa8c92893c5e/samples.tsv") %>% 
  select(sample_name, sample_catalog_tags, sample_tags) %>% 
  mutate(parse_tags = map(sample_catalog_tags, parse_catalog_tags)) %>% 
  select(-sample_catalog_tags, -sample_tags) %>% 
  unnest(parse_tags) %>% 
  pivot_wider(names_from = "tag", values_from = "value") %>% 
  mutate(day = str_extract(sample_name, "Day[0-9]*")) %>% 
  separate(sample_name, into = c("subject_from_name"), extra = "drop", remove = F) %>% 
  mutate(markers = paste(`Cluster of Differentiation (CD)`, `Other Cell Marker`, `G Protein Coupled Receptor`, sep = "_"))
df
```
```{r}
print("###Subjects"); table(df$`Subject`)
print("###Subjects"); table(df$subject_from_name)
print("###Tissue"); table(df$`Tissue Source`)
print("###Marker"); table(df$markers)
```
```{r}
table(df$`Subject`, df$markers)
```

### Yellow fever

- Yellow fever vaccine 
  - 9 individuals pre- and post-vaccination 
  - PBMCs, memory T cells, activated effector T cells, effector memory T cells
  - Project ID: `dewitt-2015-jvi`
  - Data ID: `4b7dded9-9692-428c-9c3d-23239ddd9153`
  
```{r, message=F}
df <- read_tsv("/labs/khatrilab/solomonb/immune_access/data/4b7dded9-9692-428c-9c3d-23239ddd9153/samples.tsv") %>% 
  select(sample_name, sample_catalog_tags) %>% 
  mutate(parse_tags = map(sample_catalog_tags, parse_catalog_tags)) %>% 
  select(-sample_catalog_tags) %>% 
  unnest(parse_tags) %>% 
  pivot_wider(names_from = "tag", values_from = "value")
df  
```

```{r}
table(df$Subject, df$`Time Point`, df$`Tissue Source`)
```



